Assessing United States county-level exposure for tropical cyclone impact research

### 2020-07-29 ehp_revision/response_to_reviewers.Rmd

Other approaches have also been suggested to reduce ecological bias while 
leveraging ecological data, which might be available at a much larger scale---and
so provide more power---than is practical in collecting individual data. These
include multi-level approaches in which individual-level data from a sample
of individuals in each community is incorporated.

The second issue is exposure measurement error/misclassification. When the 
exposure (e.g., binary exposure or continuous measure of intensity of exposure)
is only measured at the county level, all the individuals in that county are 
assigned that same exposure. However, the hazards included in this study will 
vary across each county. For some hazards, this variation will be more stark. 
For example, tornadoes can be very local, causing extreme damage to a row of 
homes and none to the homes across the street. Some types of floods can also 
be very localized, rather than creating widespread flooding throughout 
the county. In both cases, when a county-level exposure assessment is made
based on an event in the county in association with the storm, there may be
many individuals in the county who did not experience that hazard at their 
residence. The other hazards we included---storm-associated winds and rainfall---will
also have some variation across counties, but this will often not be as stark. 
The rainfields for tropical cyclones are very large, and while there are rainbands
within the storm that might have particularly high rates of precipitation, these
progess over the course of the storm, and it is unlikely that a county will have
one area that experienced very extreme precipitation while another experienced
very little. Winds might have more variation, especially in coastal counties and 
counties near the storm's central track, since winds decay quickly as you move
from the storm's center, and storms typically weaken rapidly once they make 
landfall and can no longer draw on warm ocean waters for energy.

Data that is aggregated across a population---for example, the total number of 
deaths in a geographic area in a certain time period---is known as *ecological
data*, and studies that base inference on this type of data are known as *ecological
studies*. Terms include *ecological*, *aggregate*, *contextual level*.

> "In an ecological study the unit of analysis is a group of people rather than 
the individual." [@sedgwick2014ecological]

> "In his seminal paper, Robinson (1950) distinguished between two types of 
correlation---ecological and individual. The former is obtained for a group
of people, while the latter is estimated for indivisible units, such as individuals."
[@portnov2007ecological]

Here is one challenge when working with aggregated data. If you use aggregated
data to infer individual-level associations, this can lead to *ecological fallacy*. 
This results from cross-level inference---the data used to model the association
is at the contextual level (e.g., county-level), while the inference is for the
individual association between exposure and outcome.

One definition of the ecological fallacy:

> "This spurious result provides an example of the ecological fallacy, in 
which conclusions ... based on ecologic data are opposite of those drawn 
on the basis of individual-level data." [@wakefield2008overcoming]

> "It is a bias produced when analyses realised in an ecological (group level)
analysis are used to make inferences at the individual level."
[@delgado2004bias]

> "Results from ecological studies are prone to the ecological fallacy. The 
ecological fallacy is a term used when collected data are analysed at a 
group level and the results are assumed to apply to associations at the 
individual level." [@sedgwick2014ecological]

Importantly, aggregated data can be used both to infer a contextual effect
(e.g., the association between county-wide rates of a health outcome and 
exposure) as well as an individual-level effect (e.g., how a person's
risk of the outcome is associated with that person's exposure to the hazard). 
The first inference will *not* be prone to ecological bias, as the inference
is at the same level as the data. The second inference (individual-level) 
will be susceptible to bias when aggregated data is used to fit the model. 
In this case, the association estimated based on aggregated data can be 
different, and even reversed, compared to the true individual-level effect.

> "I propose three criteria for the identification of ecological fallacy; 
all three of these should be present to confirm its existence: (1) Results
must be obtrained with ecological (population) data; (2) Data must be inferred
to individuals. One use of ecological studies is to explore individual-level 
association when individual data are not available. When the focus of the 
study was contextual or based on population effects and there is no inference
to individuals, ecological fallacy is not possible. When only the first two 
criteria are present---which is insufficient to affirm ecological fallacy---it
is appropriate to acknowledge that there is a possible relationship and that further 
study is required; (3) Results obtained with individual data are contradictory."
[@idrovo2011three]

> "One important breakdown in the analogy between ecologic-bias and individual-level
confounding occurs because in etiologic research, the target of inference for
both ecologic and individual-level studies is the same: Both study effects at the 
individual level. For an individual-level study, these target effects are at the 
same level as the units of analysis. But, for an ecologic study, these target 
effects are at a finer level than the units of analysis. As a result, an ecologic 
study can be unbiased for ecologic effects (in particular, ecologic confounding
may be absent) and yet still be biased for individual-level effects. Many of the 
classic social science examples of ecologic bias are of this form." [@greenland1994invited]

> "We refer to an incorrect extrapolation from unbiased estimates of ecologic effects
to unobserved individual-level effects as 'cross-level bias' (although the latter
term is sometimes used to refer to any ecologic bias). As the examples suggest, 
it is possible to detect cross-level bias if one can identify and observe 
homogeneously exposed regions." [@greenland1994invited]

> "assuming that associations observed at the level of the area hold for the 
individuals within the areas can lead to the so-called ecological fallacy 
(Selvin, 1958)." [@wakefield2006health]

> "A broad definition of ecological fallacy is that certain data (e.g., illiteracy 
rates by foreign-born) are unavailable, leading to the use of proxies and 
erroneous estimates. However, we advocate a more narrow definition of this 
phenomenon, according to which 'ecological fallacy' refers solely to differences
in conclusions which may be drawn from group-level data (e.g., average illiteracy 
rates by foreign-born), as opposed to data obtained for individuals."
[@portnov2007ecological]

> "If different conclusions are drawn from the analysis of data upon their aggregations
into units of different sizes (e.g., from individuals to townships and regions), these
differences are commonly referred to as 'ecological fallacy'. In epidemiology, these
differences are also known as ecological or cross-level bias [refs]. The ecological 
bias is difficult to control and may lead, under certain circumstances, to spurious 
effects [refs]. If epidemiological study does not consider the possibility of 
ecological bias, misguided policy decisions may follow [refs]." [@portnov2007ecological]

Ecological bias can *only* happen if there is, indeed, an individual-level effect:

> "If exposure has no effect on any individual, there will be
no individual-level or ecologic effects, and so cross-level bias cannot occur. 
Thus, cross-level bias will not affect the validity of an ecologic test of the 
null hypothesis, although it still must be considered in interpreting a 
significant effect." [@greenland1994invited]

In some cases, the ecologic-level effect (contextual effect) will be directly 
of interest, and so we may be interested in inferring at that level (in which 
case use of ecologic-level data is appropriate and will not be prone to 
ecological bias):

> "The bias phenomenon in example 7 may be ascribed to the fact that an 
ecologic (regional) variable (percent Protestant) had effects on individual 
risk, in addition to effects of the corresponding individual-level variable
(religion) that the ecological variable summarized. In both social and 
infectious disease epidemiology, as well as in community intervention studies, 
ecologic effects may be of direct interest; a classic example is the 
phenomenon of 'herd immunity', in which the overall prevalence of 
immunity in a region, as well as individual immune status, determines the 
risk of individuals within the region (16). For noncontagious diseases, 
however, ecologic effects may not be of direct interest, and may in fact
obscure the individual-level effect of interest." [@greenland1994invited]

In some cases, the estimated association based on ecological data will be very 
different from the association that exists at the individual level:

> "We see that the ecological relative risk associated with low birth 
weight is completely incomparable with the individual-level coefficient."
[@wakefield2008overcoming]

Ecological bias can also complicate estimation of effect modification, which 
otherwise could help in identifying vulnerabilities and susceptibilities among
certain subpopulations:

> "Furthermore, non-White race now appears protective, rather than detrimental
as the individual-level analysis suggests." [@wakefield2008overcoming]

Ecologic bias crops up because the aggregated data fails to capture variation
within each aggregation level (county, for example) for both the exposure 
and potential confounders. For example, during a tropical cyclone, different
parts of a county will experience different peak wind speeds. If the county 
is coastal, the most severe winds will likely be near the coast, as the 
storm will weaken once it makes landfall and moves inland, as the storm 
generates its power from the water. When we assign the entire county a
single estimate of intensity, it smooths over the fact that locations in 
some parts of the county were subject to higher peak winds while locations
in other parts of the county had lower winds. At the least, this smoothing
of variation in exposure levels across the county could lower the power of 
the study to detect a clear association between exposure and health risk, as
this smoothing drops information inherent in the variation within the county
in exposure levels.

This, however, would not result in ecological fallacy unless another factor
comes into play---variation within each aggregation level in confounders. 
Just as aggregation smooths over within-area variation in exposure levels, it
also smooths over within-area variation in levels of potential confounders. 
Depending on the patterns of this within-area variation, a result could be
that ecological-level control of the confounders does not, in fact, control
for their role at the individual level, and so the association inferred at
the ecologic level continues to be confounded by them when inferred to the
individual level.

Essentially, this comes down to a question of the joint distribution, at the
individual level, of exposure, outcome, and covariates [?].

> "As various discussants have pointed out (3--6, 8, 9), ecologic relative-risk
estimates can be subject to biases not present in estimates from individual-level
observational studies of the same populations (case-control and cohort studies).
Unlike an individual-level study, an ecologic study does not link individual
outcome events to individual exposure or covariate histories, nor does it
link individual exposure and covariate histories to one another. It is these
linkage failures that are the source of the special biases of ecologic studies
(1, 2, 6)." [@greenland1994invited]

> "In the extreme, ecologic bias in comparisons limited to homogeneously exposed
regions (as in examples 3--6) can be viewed as purely an issue of confounding,
albeit with special complexities of measurement and control of confounders."
[@greenland1994invited]

> "More generally, one may show that cross-level bias will not occur if the 
individual-level effects of all variables (including unmeasured background
factors) follow a multiple linear-regression model with no regional effects and
no interactions (17). Nevertheless, given the usual inappropriateness of the 
multiple-linear model and the absence of homogeneously exposed regions, the 
possibility of cross-level bias adds a dimension to ecologic bias beyond
that of simple confounding." [@greenland1994invited]

> "Covariate control in ecologic studies requires attention to details not 
ordinarily of concern in individual-level studies. When, as is usually the
case, important nonlinearity or nonadditivity can be expected among 
exposure and covariate effects, it may be necessary to obtain and 
control for multiple summaries of joint covariate distributions in order
to insure that control is adequate." [@greenland1994invited]

> "Ecological fallacy can be produced by within group (individual level)
biases, such as confounding, selection bias, or misclassification, and
by confounding by group or effect modification by group." [@delgado2004bias]

> "**Confounding by group**: it is produced in an ecological study, when 
the exposure prevalence of each community (group) is correlated with 
the disease risk in non-exposed of the same community. It can be 
a mechanism for producing ecological fallacy." [@delgado2004bias]

If individual-level inference is the aim, and population-level data is
available, there are some methods for using it while still aiming to avoid
ecological bias. Indeed, it can be helpful to use population-level data, as it
is often available for a large population, improving the power and precision of
the study. Further, the level of exposure might vary a lot more over the
population captured with population-level data compared to the variation that
captured in a smaller sample of individual-level data. This can contribute both 
to statistical power and improve external validity (as the study data will cover
more of the range of exposure that might ever be expected). However, if
individual-level inference is the goal, then there are ways to supplement
population-level data with samples of individual-level data through two-level, 
semi-ecologic study designs.

> "It is generally well recognized that in order for ecologic data to provide
reliable inferences, they need to be supplemented with individual-level data."
[@wakefield2008overcoming]

> "For this example, we have access to complete individual-level data, permitting
a 'gold standard' individual-level analysis." [@wakefield2008overcoming]

> "The only reliable way to characterize within-area variation in exposures
and confounders, and hence control ecologic bias, is to collect and incorporate
individual-level data. To help epidemiologists achieve this goal, in this paper
we describe the use of the two-phase design in an ecologic setting."
[@wakefield2008overcoming]

> "The only reliable way to characterize within-area variation in exposures
and confounders, and hence control ecologic bias, is to collect and incorporate
individual-level data. To help epidemiologists achieve this goal, in this 
paper we describe the use of the two-phase design in an ecologic setting."
[@wakefield2008overcoming]

> "In a semi-ecologic study, an ecologic exposure is combined with individual-level
outcomes and confounders. A two-phase approach is particularly useful for
such a study, with the phase 2 data corresponding to stratified sampling of 
individual exposures." [@wakefield2008overcoming]

> "To implement a two-phase design, an initial phase 1 cross-classification by 
the binary disease outcome and stratification variables is required; in phase
2, samples of individuals are drawn from each of the cross-classification cells, 
with data on additional variables being drawn from the subsamples of individuals. 
Intuitively, the stratified sampling is focused on informative cells, and 
estimation methods use both phases of data for efficiency and to acknowledge
the outcome-dependent sampling. In the simplest ecologic setting, the cross-classification
is by outcome and area only, and if area is a surrogate for important risk factors
this design will be efficient. We are particularly interested in situations where
an initial classification is available by outcome, area, and confounders such as
age and gender---this is the case in a semi-ecologic study. Phase 2 may then provide
detailed exposure information on a subset of the phase 1 individuals." 
[@wakefield2008overcoming]

> "Two-phase study designs are a generalization of matched case-control designs
in which, initially, the entire sample population is cross-classified according
to case/control status and some stratification variable, S. The latter depends
on covariates observed in all individuals and may include exposures of interest, 
proxy exposure measures, or potential confounders. In settings like those we
are considering, such as environmental epidemiology, S may also depend on geographic
area, which can act as a surrogate for the totality of confounders associated
with each area, as well as provide a well-defined sampling frame for the controls."
[@wakefield2008overcoming]

> "The results of our simulation study point to the benefits associated with combining
the two sources of data, in terms of both bias and efficiency. Rather than supplement
an ecologic study with individual-level data, it may be of interest to combine 
existing individual-level data with external group-level data. Strategies that
combine both types of data have been shown to alleviate participation bias and
improve efficiency in case-control studies with missing data (31)." 
[@wakefield2008overcoming]

> "As with other observational studies, ecologic studies can give useful results if 
biases such as those discussed here can be ruled out or quantified. Nevertheless, 
bias evaluation can be especially difficult in ecologic studies of geographic
regions because of the many potentially interacting covariates that may differ
across regions. When biases cannot be ruled out with available data, further
exploration will require individual-level studies." [@greenland1994invited]

Why we might sometimes use (and want to use) population-level data:

> "Because of the unavailability of individual-level data, ecological data
may be resorted to and may come in a variety of forms." [@wakefield2008overcoming]

> "Epidemiologists continue to use ecologic and aggregate data. Despite their 
known drawbacks, these data, often aggregated across geographic areas, offer the
advantages of widespread availability and gains in statistical power from large
populations and increased exposure ranges." [@wakefield2008overcoming]

> "Population and health data are often routinely available in ecological, 
that is group, form while the exposure data typically consist of a set of
values recorded at monitor sites, or via one-off sampling." [@wakefield2006health]

It can be important to think about the scale that the process happens at. For
something very local, you might lose a lot more with aggregated data.

> "Data availability and exposure variability often determine the scale of
examination and the suitability of a study. Exposures arising from a point or
line source offer exposure contrasts on small scales, requiring small-area data;
in constrast, dietary variables show little variation across small scales, and
consequently international studies are used." [@wakefield2008overcoming]

There can also be an *individualistic fallacy*, if emphasis is limited to the
association between individual-level factors, without considering the potential
role of contextual factors in an individual's risk of a health outcome. In other
words, cross-level bias can go in both directions [@idrovo2011three]. While the
fallacy of inferring individual-level associations from ecologic-level data is
called *ecological fallacy*, the fallacy of inferring ecologic-level
(contextual) effects from individual-level data is known as the *atomistic
fallacy* [@idrovo2011three]. There are also fallacies that can come in, it
seems, from inferring at the same level as the data but failing to consider
influences from factors at the other level. When inferring individual-level
associations from individual-level data, without considering an additional role
of ecological-level factors, this is known as the *psychologistic* or
*individualistic fallacy* [@idrovo2011three]. When inferring contextual effects
from ecologic data without considering individual-level factors, this is known
as the *sociologistic fallacy* [@idrovo2011three].

> "Although ecological studies are important to epidemiology (especially in 
environmental and social epidemiology), public health practitioners seem 
afraid of ecological studies. It is a common practice to assume the presence of 
ecological fallacy (Robinson 1950) and low-level validity when analyzing an 
ecological study. Most epidemiologists prefer an exclusive individualistic
approach, although the importance of a multilevel causal approach is widely 
recognized (Diez-Roux 2002). In this sense, some authors suggest that it is
as important to recognize the presence of ecological fallacy as to 
recognize psychologistic or individualistic fallacy (Subramanian et al. 2009)."
[@idrovo2011three]

### 2020-07-29 ehp_revision/response_to_reviewers.Rmd

> **Text in revised manuscript:**
"While other approaches have been developed to incorporate storm
hazards, particularly wind, into exposure assessment (e.g., Grabich et al.
2015a; Zandbergen 2009; Czajkowski et al. 2011), there is not yet a standard
approach. When different studies use different datasets or storm hazards
when assessing storm exposure, it becomes difficult to compare and aggregate
findings."

\begin{shaded}
\textbf{R1 Comment 6:}
This is not a comment on the paper, but the authors may also want to consider
publishing their work in DesignSafe, which is "the web-based research platform
of the NHERI  [NSF-Funded Natural Hazards Engineering Research Infrastructure]
Network that provides the computational tools needed to manage, analyze, and
understand critical data for natural hazards research." It is the data
repository widely used by interdisciplinary hazards and disaster researchers.
\end{shaded}

**Response:**

Great idea!

We have investigated this. DesignSafe is ...

As part of its platform, DesignSafe allows users to work in a number of 
programming languages, including R (through Jupyter Notebooks). Therefore, 
DesignSafe users already have access to any data shared through an R package, 
including the data we present in this manuscript, without the data needing 
to be separately loaded and published through DesignSafe's Data Depot.

We have added a note in the manuscript to let readers know that, because the
data is shared through an R package, and because DesignSafe's [Workspace]
allows users to work in R, the data can be fully accessed and used for those working
on the DesignSafe platform by using the same code in a DesignSafe Jupyter notebook
that would otherwise be used on a local computer.

This is good, because otherwise there would be a good deal of overlap between
the data on DesignSafe and that in the R package. R packages already created 
a citable source, comparable to the DOI assigned to data by DesignSafe when
it is published. Further, R packages include licensing information, as does
DesignSafe for published data. Finally, if the data were shared through both 
platforms, it would be critical to make sure that it was clear that the two 
datasets are analogous and also to keep them up to date with each other, which
would add an extra layer of work with each update to the R package (we have 
been updating several hazards as new data becomes available). It is wonderful 
that DesignSafe's interface with R avoids this problem by allowing its users
direct access to our package and its data through that path, as it allows for 
a single, citable, licensed version of the data stored and shared from a single
location, with no question of syncing separate versions of the data or clarifying
licensing and citation practices across two storage locations.

*What is DesignSafe:*

> "DesignSafe is the cyberinfrastructure platform that has been developed as
part of NHERI to support natural hazards engineering research, and
it succeeds the NEEShub cyberinfrastructure that was developed for the
earthquake engineering community through the Network for Earthquake
Engineering Simulation (NEES) program." [@rathje2017designsafe]

> "DesignSafe plays an important role in integrating the various NHERI components
and the research taking place at NHERI facilities, but also has the broader
goal of enabling transformative research in natural hazards and engineering
across the numerous technical disciplines engaged in this field." 
[@rathje2017designsafe]

> "DesignSafe has been developed as a flexible, extensible, community-driven
cyberinfrastructure, and it embraces a cloud strategy for the big data generated
in natural hazards engineering. DesignSafe provides a comprehensive CI that
supports the full research lifecycle from planning to execution to analysis
to publication and curation." [@rathje2017designsafe]

> "The future of natural hazards engineering research requires the integration
of diverse data sets from a variety of sources including experiments, computational
simulation, and field reconnaissance, as well as a variety of research disciplines
including earth science, social science, building science, and architecture. 
The DesignSafe cyberinfrastructure has been designed to provide the functionalities
that will enable transformative research in natural hazards engineering. By adopting
a cloud strategy, DesignSafe allows for a fundamental change in the way that
research is performed. It provides a comprehensive cyberinfrastructure that
supports research workflows, data analysis, and visualization, as well as the 
full lifecycle of experimental, field, and computational research required by 
engineers and scientists to effectively address the threats posed to civil 
infrastructure by natural hazards. The integration of data and computation 
in the cloud will enable new research discoveries in natural hazards engineering, 
which in turn can lead to more hazard-resilient civil infrastructure." [@rathje2017designsafe]

*Access to R:*

> "The real revolution takes place downstream of data generation. Here, data 
analytics and visualization are performed in the cloud within the *Discovery
Workspace* while accessing any data within the *Data Depot*. Researchers can 
invoke common analysis programs, such as MATLAB, as well as other analysis/
visualization tools, such as Jupyter notebooks. A Jupyter notebook is an 
electronic notebook that allows users to embed rich text elements as well
as computer code, graphs, and visualizations within a single notebook that
can be shared through the web. Over 40 different programming languages are 
supported in Jupyter, including Python and R, and MATLAB code can be easily 
converted, making Jupyter a versatile tool for research. Performing analysis in 
the cloud allows researchers to integrate and explore various data without
tedious downloads. Additionally, using a seamless cyberinfrastructure to 
complete all research tasks enables tracking and relating of the processes 
applied to data." [@rathje2017designsafe]

*Data Depot:*

> "the Data Depot, a flexible data repository with streamlined data management
tools." [@rathje2017designsafe]

> "As data progresses towards publication the requirements for metadata 
increase as metadata provides users with search and discovery functions. At
the end of the research project the user may edit the information for 
publication and complete the process of assigning digital object identifiers
(DOIs) and applying the appropriate license. On-demand assistance from a 
curator is available to provide training and guide users through their data
curation and publication needs." [@rathje2017designsafe]
